An application of Graph Theory to MDAO problem formulation

### 2012-10-02 intro.tex
updated the simple problem to use better icons

\section{Fundamental Problem Formulation}
    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem with two analyses
    the problem formulation is as follows

    \begin{align}
        given & \ \ A.f\left(A.x,A.y\right) \notag
        \\ & \ \  B.f\left(B.x,B.z\right) \notag
        \\ & \ \  C.f1\left(A.f,B.f,C.s\right) \notag
        \\ & \ \  C.f2\left(A.f,B.f,C.s\right) \notag
        \\min &\ \ F\left(C.f1\right) \notag
        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
        \\      &\ \ E\left(A.y,C.f1\right) = 0 \notag
        \\      &\ \ R\left(C.s,C.f2\right) = 0 
        \label{eqn:simple}
    \end{align}

    Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
    has a set of input and output variables. While this does represent a complete problem formulation, 
    it's not the fundamental one since some assumptions have inherently been made about the solution 
    strategy. $C$ is directly dependent
    $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and
    iterate on the design variables in $A$ and $B$ until you satisfy $E$. However, a different solution could be equally valid and still represent
    the same fundamental problem:

    \begin{align}
        given & \ \ A.f\left(A.x,C.f1\right) \notag
        \\ & \ \  B.f\left(B.x,B.z\right) \notag
        \\ & \ \  C.f1\left(C.y,B.f,C.s\right) \notag
        \\ & \ \  C.f2\left(C.q,B.f,C.s\right) \notag
        \\min &\ \ F\left(C.f1\right) \notag
        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
        \\      &\ \ E\left(A.f,C.q\right) = 0 \notag
        \\      &\ \ R\left(C.s,C.f2\right) = 0 
        \label{eqn:simple_2}
    \end{align}

    Equation \ref{eqn:simple_2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now directly dependent on the output of $C$, 
    and the constraint $E$ has changed as well. Now the problem must be solved by running $B$ and $C$, then $A$ and iterating 
    on the design variables in $B$ and $C$ until you satisfy $E$. Since the formulations in Eqn. \ref{eqn:simple} 
    and \ref{eqn:simple_2} both describe the same problem there must be a more fundamental description of the problem that is common
    between them. We present the FPF as follows:

    \begin{align}
        given & \ \ A.f\left(A.x,A.y\right) \notag
        \\ & \ \  B.f\left(B.x,B.z\right) \notag
        \\ & \ \  C.f1\left(C.y,C.r,C.s\right) \notag
        \\ & \ \  C.f2\left(C.q,C.r,C.s\right) \notag
        \\min &\ \ F\left(C.f1\right) \notag
        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
        \\      &\ \ E\left(A.f,C.\right) = 0 \notag
        \\      &\ \ R\left(C.s,C.f2\right) = 0 
        \label{eqn:simple_2}
    \end{align}

>>>>

\section{Fundamental Problem Formulation}
    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
    the problem formulation is as follows:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z}
              & \ \ B: {x,z} \rightarrow {y^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G(y^t,y) = 0
        \label{eqn:simple}
    \end{align}

    Where $A$ and $B$ represent analysis tools, $F,G$ are the objective and constraint functions respectively. 
    While this does represent a complete problem formulation, 
    it's not the fundamental one since some assumptions have inherently been made about the solution 
    strategy. Analysis $A$ outputs $z$ which is an input to $B$. Hence, $A$ should be run before $B$ with 
    the $y$ being iterated on to convergence. However, a different solution could be equally valid and still represent
    the same fundamental problem:

    \begin{align}
        given & \ \ B: {x,z} \rightarrow {y}
              & \ \ A: {x,y} \rightarrow {z^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G(z^t,z) = 0
        \label{eqn:simple2}
    \end{align}

    Equation \ref{eqn:simple2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
    and $z$ will be iterated on to meet $G$. Now the problem can be solved by running $B$ first and then $A$.
    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple_2} both describe the same problem 
    there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z^t}
              & \ \ B: {x,z} \rightarrow {y^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G1(z^t,z) = 0 \notag
             & \ \ G2(y^t,y) = 0
        \label{eqn:simple_fpf}
    \end{align}

    The FPF in Eqn. \ref{eqn:simple_fpf} differs from both Eqn. \ref{eqn:simple} and Eqn. \ref{eqn:simple2} because it has 
    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint, and replacing 
    it with a direct dependence between the two analyses you could regain the earlier two problem formulations.
